---
title: "PMM3: Estimation for Symmetric Platykurtic Errors"
author: "Serhii Zabolotnii"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PMM3: Estimation for Symmetric Platykurtic Errors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

## When to Use PMM3 Instead of PMM2 or OLS

The Polynomial Maximization Method of order 3 (PMM3) is designed for situations
where regression errors are **symmetric** but **platykurtic** (lighter tails
than the normal distribution). Common examples include uniform, beta-symmetric,
and truncated normal errors.

| Error property          | Recommended method |
|-------------------------|--------------------|
| Skewed (|gamma3| > 0.3) | PMM2 (`lm_pmm2`)  |
| Symmetric, platykurtic (gamma4 < -0.7) | **PMM3** (`lm_pmm3`) |
| Near-Gaussian           | OLS (`lm`)         |

You can use `pmm_dispatch()` to automatically determine the best method:

```{r dispatch-example}
library(EstemPMM)

set.seed(42)
n <- 200
x <- rnorm(n)
eps <- runif(n, -sqrt(3), sqrt(3))  # Uniform: symmetric, platykurtic
y <- 1 + 2 * x + eps

fit_ols <- lm(y ~ x)
recommendation <- pmm_dispatch(residuals(fit_ols))
```

## Basic Usage

```{r basic-fit}
fit_pmm3 <- lm_pmm3(y ~ x, data = data.frame(y = y, x = x))
summary(fit_pmm3)
```

## Interpreting the Results

PMM3 reports several key quantities:

- **m2, m4, m6**: Central moments of initial (OLS) residuals
- **gamma4**: Excess kurtosis (negative = platykurtic)
- **gamma6**: Sixth-order cumulant coefficient
- **g3**: Theoretical efficiency factor — `Var(PMM3) = g3 * Var(OLS)`
- **kappa**: Moment ratio used internally by the Newton-Raphson solver

A smaller `g3` means greater variance reduction over OLS.

## Comparison with OLS

```{r ols-comparison}
coef_ols  <- coef(fit_ols)
coef_pmm3 <- coef(fit_pmm3)

comparison <- data.frame(
  OLS  = round(coef_ols, 5),
  PMM3 = round(coef_pmm3, 5),
  Diff = round(coef_pmm3 - coef_ols, 5)
)
print(comparison)

cat("\nResidual sum of squares:\n")
cat("  OLS:  ", sum(residuals(fit_ols)^2), "\n")
cat("  PMM3: ", sum(residuals(fit_pmm3)^2), "\n")
```

## Adaptive Mode

By default, PMM3 computes kappa from OLS residuals and holds it fixed. In
adaptive mode, kappa is re-estimated from the current residuals at each
Newton-Raphson iteration:

```{r adaptive}
fit_adaptive <- lm_pmm3(y ~ x, data = data.frame(y = y, x = x), adaptive = TRUE)
cat("Fixed kappa coefficients: ", round(coef(fit_pmm3), 5), "\n")
cat("Adaptive coefficients:    ", round(coef(fit_adaptive), 5), "\n")
```

## Diagnostic Plots

```{r diagnostics, fig.height=6}
plot(fit_pmm3)
```

## Time Series Models

PMM3 supports AR, MA, ARMA, and ARIMA models through dedicated functions.

### AR Model

```{r ar-model}
set.seed(42)
innov <- runif(200, -sqrt(3), sqrt(3))
ar_data <- arima.sim(n = 200, list(ar = 0.7), innov = innov)

fit_ar <- ar_pmm3(ar_data, order = 1)
summary(fit_ar)
```

### ARMA Model

```{r arma-model}
innov2 <- runif(250, -sqrt(3), sqrt(3))
arma_data <- arima.sim(n = 250, list(ar = 0.5, ma = 0.3), innov = innov2)
fit_arma <- arma_pmm3(arma_data, order = c(1, 1))
coef(fit_arma)
```

### Forecasting

```{r forecast}
predict(fit_ar, n.ahead = 5)
```

## Utility Functions

### Testing Symmetry

```{r symmetry}
test_symmetry(residuals(fit_ols))
```

### Sixth-Order Cumulant

```{r gamma6}
pmm_gamma6(residuals(fit_ols))
```

### Variance Factor

```{r variance-factor}
mom <- compute_moments_pmm3(residuals(fit_ols))
pmm3_variance_factor(mom$m2, mom$m4, mom$m6)
```
